---
title: "TDaun_Livesession05assignment"
author: "Travis Daun"
date: "12/9/2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1. Data Munging
### a. Import Data
```{r}
# Import the data
df <- read.table("datasets/yob2016.txt")
str(df)
# From str function can determine the data is delimited by ;
df <- read.table("datasets/yob2016.txt", sep=";")
str(df)
# Make human readable column names
colnames(df) <- c("firstName", "gender", "amountOfChildren")
# Cast the factor variables to character variables
df$firstName <- as.character(df$firstName)
df$gender <- as.character(df$gender)
```
### b. Display the summary and structure
```{r}
#Summary of data
summary(df)
#Structure of Data
str(df)
```
### c. Figure out which name has yyy
```{r}
#Search the end of the string for "yyy"
grep("yyy$", df$firstName)
#Examine the data around that to verify it is a duplicate
df[210:214,]
```
### d. Remove this particular observation and save as y2016
```{r}
# Remove duplicate(misspelled) row
y2016 <- df[-c(212),]
# Verify it was removed
y2016[210:214,]
```

## 2. Data Merging
### a. Import 2015 data
```{r}
# Import the data
y2015 <- read.table("datasets/yob2015.txt")
str(y2015)
# From str function can determine the data is delimited by ,
y2015 <- read.table("datasets/yob2015.txt", sep=",")
str(y2015)
colnames(y2015) <- c("firstName", "gender", "amountOfChildren")
# Cast the factor variables to character variables
y2015$firstName <- as.character(y2015$firstName)
y2015$gender <- as.character(y2015$gender)
#Summary of data
summary(y2015)
#Structure of Data
str(y2015)
```
### b. Display the last ten rows in the dataframe. Describe something you find interesting about these 10 rows.
* Zyrus and Zyus could be the same name with a misspelling. Could be be duplication of data.
```{r}
# Display last 10 rows of dataframe
tail(y2015, 10)
```
### c. Merge y2016 and y2015 by your Name column; assign it to final.
```{r}
final <- merge(y2015, y2016, by = "firstName")

# Remove any rows that did not have names in 2015 and 2016
final <- subset(final, final$amountOfChildren.x >=1 & final$amountOfChildren.y >=1)

# Remove the rows where gender did not match
final <- subset(final, final$gender.y == final$gender.x)

# Remove the duplicate gender column
final$gender.y <- NULL

str(final)
colnames(final) <- c("firstName", "gender", "amountOfChildren2015", "amountOfChildren2016")
```

## 3. Data Summary
### a. Total 2015 and 2016 number of children with the same name.
```{r}
# Total the 2015 and 2016 children
final$total = final$amountOfChildren2015 + final$amountOfChildren2016
```

### b. Sort the data by Total. What are the top 10 most popular names?
```{r}
# Sort in decending order
final <- final[order(-final$total),]
# Top 10 most popular names
head(final, 10)
```
### c. Omit boys and give the top 10 most popular girl’s names.
```{r}
# Omit boys names
final <- subset(final, final$gender == "F")
head(data.frame(final$firstName, final$total), 10)
```
### d. Write these top 10 girl names and their Totals to a CSV file
```{r}
write.csv(head(data.frame(final$firstName, final$total), 10), file = "girlsNames.csv", row.names = FALSE)
```

## 4. Upload to GitHub
### https://github.com/dauntr/SMU_MSDS_Homework